Refinement of Thermostated Molecular Dynamics via Backward Error Analysis

### 2018-06-16 rigid_bodies.tex
Some small changes

\documentclass[
	aip,
	jcp,
%	preprint,
	reprint,
]{revtex4-1}

@@

\subsection{NVT Dynamics with a Single Thermostat Chain: Notation and Measure-Preserving Integration}
\label{sec:nvt}

@@

Despite the non-equilibrium process occurring within the thermostats, it is worth showing the results obtained with the method of Kamberaj \textit{et al}. \cite{Kamberaj_2005}, because that is not the only problematic issue.
Particularly, from Fig.~\ref{fig:properties}(a) we note that the simulated temperature decreases systematically with $\timestep$.
This behavior is observed in NVE simulations\cite{Davidchack_2010, Silveira_2017}, reflecting that the kinetic energy is weakly controlled in that factorization scheme.
As a result, the potential energy and virial are not severely influenced by $\timestep$ as they are in the scheme of Martyna \textit{et al}. \cite{Martyna_1996}. Figs.~\ref{fig:properties}(a) and \ref{fig:properties}(b) clearly show that the temperature control described in Sec.~\ref{sec:nvt} leads to a severely degraded accuracy in $\langle U \rangle$, which systematically increases with $\timestep$.
As demonstrated in Fig.~\ref{fig:nve}, the discretization errors are non-negligible even for a small step size of 2 fs.
Thus, the results for the method of Martyna \textit{et al}. \cite{Martyna_1996} can be interpreted as follows: if the kinetic energy is restricted to adopt a certain value, in accordance with the target temperature, the unavoidable discretization errors will accumulate in the remaining degrees of freedom.

From Fig.~\ref{fig:properties} it becomes clear that the refined schemes yield significantly improved results: we are able to simultaneously reproduce the specified temperature and the other thermodynamic properties.
Note that a small, still clearly influencing systematic error is present, as the computed averages are slightly modified for $\timestep$ > 5.
This could, in principle, be remediated by using a higher-order approximation to the shadow Hamiltonian.
Mathematically, the derivation of the refined Hamiltonian was facilitated by the fewer number of operators associated with the unsplit solution of rotation.
However, the accuracy of the approximate solution is very high for the time step sizes considered\cite{Silveira_2017} and this explains the good agreement between schemes Refined/Unsplit and Refined/NO-SQUISH.

@@

With the aim of highlighting the importance of reweighting, in Fig.~\ref{fig:free_energy} we show for each time-step size the free energy difference per water molecule ($\Delta F/N$, circles) between the sampled and target states.
$\Delta F$ corresponds to the logarithm of the averaged denominator of Eq.~\eqref{eq:reweighting}.
As expected, the free energy is only negligible for $h$ = 1 fs.

@@

We end this section with a discussion of discretization effects on the kinetic energy partition.
In Fig.~\ref{fig:energy_partition} we show the translational ($K_t$) and rotational ($K_r$) kinetic energies as a function of the time-step size, and compare the results with the expected values assuming equipartition.
In Figs.~\ref{fig:energy_partition} (a)-(c) these were calculated from the kinetic energy $\langle K \rangle$, as follows: $\langle K_t \rangle_{eq}$ = $\frac{\langle K \rangle}{6N-3} \times (3N-3)$ and $\langle K_r \rangle_{eq}$ = $\frac{\langle K \rangle}{6N-3} \times 3N$.
For Fig.~\ref{fig:energy_partition} (d) we substituted $\langle K \rangle$ for $\langle \refined{K} \rangle$ in the right-hand side of the equations above.

In our previous work\cite{Silveira_2017}, we demonstrated that in the NVE case the rotational kinetic energy is more influenced by the time-step size because the rotational degrees of freedom are the fastest modes in the rigid model of water. 
We see that the method of Kamberaj \textit{et al}. \cite{Kamberaj_2005} reproduces the behavior of the NVE case, as noted before.
The results of Fig.~\ref{fig:energy_partition} (a), which were obtained for the numerical scheme of Martyna \textit{et al}. \cite{Martyna_1996}, are similar to Fig.~\ref{fig:energy_partition} (b), in the sense that the simulated translational and rotational kinetic energies deviate almost symmetrically and in opposite directions.
Also, they do so contrary to what is expected: the translational kinetic energy should be smaller than the rotational one.
From Fig.~\ref{fig:energy_partition} (d) one can note that the refined translational and rotational kinetic energies obey the energy partition predicted by Tolman's theorem \cite{Tolman_1918} for time-step sizes up to 5 fs.
For this range of $\timestep$ and after reweighting, which yields Fig.~\ref{fig:energy_partition} (b), it becomes clear that our method corrects the serious disruption observed in the schemes of Martyna \textit{et al}. \cite{Martyna_1996} and  Kamberaj \textit{et al}. \cite{Kamberaj_2005}.

>>>>

\documentclass[
	aip,
	jcp,
	preprint,
%	reprint,
]{revtex4-1}

@@

\subsection{NVT Dynamics with a Single Nos\'{e}-Hoover Chain: Notation and Measure-Preserving Integration}
\label{sec:nvt}

@@

Despite the non-equilibrium process occurring within the thermostats, it is worth showing the results obtained with the method of Kamberaj \textit{et al}. \cite{Kamberaj_2005}, because that is not the only problematic issue.
Particularly, from Fig.~\ref{fig:properties}(a) we note that the simulated temperature decreases systematically with $\timestep$.
This behavior is observed in NVE simulations\cite{Davidchack_2010, Silveira_2017}, reflecting that the kinetic energy is weakly controlled in that factorization scheme.
As a result, the potential energy and virial are not severely influenced by $\timestep$ as they are in the scheme of Martyna \textit{et al}. \cite{Martyna_1996}. Figs.~\ref{fig:properties}(a) and \ref{fig:properties}(b) clearly show that the temperature control described in Sec.~\ref{sec:nvt} leads to a severely degraded accuracy in $\langle U \rangle$, which systematically increases with $\timestep$.
As demonstrated in Fig.~\ref{fig:nve}, the discretization errors are non-negligible even for a small step size of $2~\text{fs}$.
Thus, the results for the method of Martyna \textit{et al}. \cite{Martyna_1996} can be interpreted as follows: if the kinetic energy is restricted to adopt a certain value, in accordance with the target temperature, the unavoidable discretization errors will accumulate in the remaining degrees of freedom.

From Fig.~\ref{fig:properties} it becomes clear that the refined schemes yield significantly improved results: we are able to simultaneously reproduce the specified temperature and the other thermodynamic properties.
Note that a small, still clearly influencing systematic error is present, as the computed averages are slightly modified for $\timestep > 5~\text{fs}$.
This could, in principle, be remediated by using a higher-order approximation to the shadow Hamiltonian.
Mathematically, the derivation of the refined Hamiltonian was facilitated by the fewer number of operators associated with the unsplit solution of rotation.
However, the accuracy of the approximate solution is very high for the time step sizes considered\cite{Silveira_2017} and this explains the good agreement between schemes Refined/Unsplit and Refined/NO-SQUISH.

@@

With the aim of highlighting the importance of reweighting, in Fig.~\ref{fig:free_energy} we show for each time-step size the free energy difference per water molecule ($\Delta F/N$, circles) between the sampled and target states.
$\Delta F$ corresponds to the logarithm of the averaged denominator of Eq.~\eqref{eq:reweighting}.
As expected, the free energy is only negligible for $h = 1~\text{fs}$.

@@

We end this section with a discussion of discretization effects on the kinetic energy partition.
In Fig.~\ref{fig:energy_partition} we show the translational ($K_t$) and rotational ($K_r$) kinetic energies as a function of the time-step size, and compare the results with the expected values assuming equipartition.
In Figs.~\ref{fig:energy_partition}(a) to \ref{fig:energy_partition}(c) these were calculated from the kinetic energy $\langle K \rangle$, as follows: $\langle K_t \rangle_\text{eq}$ = $\frac{\langle K \rangle}{6N-3} \times (3N-3)$ and $\langle K_r \rangle_\text{eq}$ = $\frac{\langle K \rangle}{6N-3} \times 3N$.
For Fig.~\ref{fig:energy_partition}(d) we substituted $\langle K \rangle$ for $\langle \refined{K} \rangle$ in the right-hand side of the equations above.

In our previous work \cite{Silveira_2017}, we demonstrated that in the NVE case the rotational kinetic energy is more influenced by the time-step size because the rotational degrees of freedom are the fastest modes in the rigid model of water. 
We see that the method of Kamberaj \textit{et al}. \cite{Kamberaj_2005} reproduces the behavior of the NVE case, as noted before.
The results of Fig.~\ref{fig:energy_partition}(a), which were obtained for the numerical scheme of Martyna \textit{et al}. \cite{Martyna_1996}, are similar to Fig.~\ref{fig:energy_partition}(b), in the sense that the simulated translational and rotational kinetic energies deviate almost symmetrically and in opposite directions.
Also, they do so contrary to what is expected: the translational kinetic energy should be smaller than the rotational one.
From Fig.~\ref{fig:energy_partition}(d) one can note that the refined translational and rotational kinetic energies obey the energy partition predicted by Tolman's theorem \cite{Tolman_1918} for time-step sizes up to $5~\text{fs}$.
For this range of $\timestep$ and after reweighting, which yields Fig.~\ref{fig:energy_partition}(b), it becomes clear that our method corrects the serious disruption observed in the schemes of Martyna \textit{et al}. \cite{Martyna_1996} and  Kamberaj \textit{et al}. \cite{Kamberaj_2005}.
